---
title: "Sup Figure 2C"
author: "Marta Milo"
date: "03/08/2023"
output: html_document
---

## Produce figure to answer review comment

Analyse the Tab3 from supplementary data. 

1. Take all synergistic combos with targets and pathways as per the threshold 0.1 HSA, 0.5 Emax 'n'
  

2. Calculate the total number of combos per pathways to full data matrix as in fig 1c 'Nc'
  Taking number of synergistic combos 'n' , and generate random numbers (from 1-n) assigned to each pathway combo
  Bootstrap 10 fold with a upper limit of 'n' and calculate average for each pathway combo category (nb)    

3. Generate 2 figs similar to 1c

Pathway combos for only syn combos evaluate the number for each category 'nc'

Pathway combos with syn vs total ratio 'Es = nc/Nc' for each pathway category - Es (enrichment for synergy)

Pathway combos with weight from bootstrap 'Er = nb/Nc' for each pathway category - Er (enrichment by random)
  Colour blob in different colour if Es > Er


## Loading the data 
```{r}
all_data<- read.csv("Supplementary Table 3. Single and combination metrics.csv", header=TRUE)

### set libraries
library(ggplot2)
library(cowplot)
library(dplyr)

library(ggpubr)
library(colorspace)

library(tidyverse)
library(gprofiler2)
library(tibble)
library(reshape2)

library(pheatmap)


```
## Calculate the number of combos for full data matrix 
```{r}

## create empty columns

all_data$LIB1 = ""
all_data$LIB2 = ""
all_data$lib2_pathway<-gsub("Other", "Chemotherapy", all_data$Compound2_pathway)
all_data$Compound1_pathway<-as.character(all_data$Compound1_pathway)
all_data$Compound2_pathway<-as.character(all_data$Compound2_pathway)

for (i in 1:nrow(all_data)){
  #Display pathway pairings in alphabetical order, i.e. only gives us the top triangle of the matrix
  all_data$LIB1[i]<-pmin(all_data$Compound1_pathway[i], all_data$Compound2_pathway[i])
  all_data$LIB2[i]<-pmax(all_data$Compound1_pathway[i], all_data$Compound2_pathway[i])
}




## grouping for pathways and defining the ns
#Note - this is not the same as 1C as 1C only contains drug combo info. The following analysis contains drug combo-cell line pairs for comparison with synergy calls

all_data_table<-table(all_data$LIB1, all_data$LIB2)
all_data_table<-as.data.frame(all_data_table)
names(all_data_table)<-c("lib1_pathway", "lib2_pathway", "Frequency")

ggplot(all_data_table, aes(lib1_pathway, lib2_pathway)) + geom_point(aes(size=Frequency))+ theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(size = "Number of cell line-combinations screened", x= "Pathway 1", y="Pathway 2") +scale_size_area( max_size = 10)

#Table of number of pathway combinations
all_data_table
```




## Filter for synergistic combos

Threshold used
HSA> 0.1
Emax >0.5

```{r}
tx<- which(all_data$combo_MaxE>0.5 & all_data$HSA_matrix>0.1, arr.ind = TRUE)

filtered_data<-all_data[tx,] 

filtered_data_table<-table(filtered_data$LIB1, filtered_data$LIB2)
filtered_data_table<-as.data.frame(filtered_data_table)
names(filtered_data_table)<-c("lib1_pathway", "lib2_pathway", "Frequency")

ggplot(filtered_data_table, aes(lib1_pathway, lib2_pathway)) + geom_point(aes(size=Frequency))+ theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(size = "Number of active and synergistic \ncell line-combinations", x= "Pathway 1", y="Pathway 2") +scale_size_area( max_size = 10)

#Table of number of pathway combinations
filtered_data_table

```

## Randomising HSA and combo Emax values, and refiltering to synergistic combos 

```{r}
#Randomly scramble HSA_matrix and  
set.seed(13)
randomised_data<-all_data
randomised_data$HSA_matrix<-sample(randomised_data$HSA_matrix, replace = FALSE)
randomised_data$combo_MaxE<-sample(randomised_data$combo_MaxE, replace = FALSE)

tx2<- which(randomised_data$combo_MaxE>0.5 & randomised_data$HSA_matrix>0.1, arr.ind = TRUE)

randomised_filtered_data<-randomised_data[tx2,] 

randomised_filtered_data_table<-table(randomised_filtered_data$LIB1, randomised_filtered_data$LIB2)
randomised_filtered_data_table<-as.data.frame(randomised_filtered_data_table)
names(randomised_filtered_data_table)<-c("lib1_pathway", "lib2_pathway", "Frequency")

ggplot(randomised_filtered_data_table, aes(lib1_pathway, lib2_pathway)) + geom_point(aes(size=Frequency))+ theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(size = "Number of active and synergistic \nrandomised cell line-combinations", x= "Pathway 1", y="Pathway 2") +scale_size_area( max_size = 10)

#Table of number of pathway combinations
randomised_filtered_data_table

```

## define enrichment for pathways over randomness

1. Pathway combos with syn vs total ratio 'Es = ncombo/Ntotal' for each pathway category - Es (enrichment for synergy)
2. Pathway combos with weight from bootstrap 'Er = nrand/Ntotal' for each pathway category - Er (enrichment by random)
3. Colour blob in different colour if Es > Er


```{r}

temp<- all_data_table
temp[,3]<- filtered_data_table[,3]/all_data_table[,3]
Es<-temp
## remove the NaN and substitute with 0 in list
is.nan.data.frame <- function(x)
do.call(cbind, lapply(x, is.nan))
Es[is.nan(Es)] <- 0


temp<- all_data_table
temp[,3]<- randomised_filtered_data_table[,3]/all_data_table[,3]
Er<-temp
## remove the NaN and substitute with 0
Er[is.nan(Er)] <- 0


## merging the data for plotting 

full_Er_Es<- cbind(Es,Er[,3])
colnames(full_Er_Es)[3:4]<-c("Es","Er")

### drowing teh matrix with enrichment score
ggplot(full_Er_Es, aes(lib1_pathway, lib2_pathway)) + geom_point(aes(size=Es, colour = (Es > Er)))+
  scale_colour_manual(values =c( "black","blue")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(size = "Number of active and synergistic \nrandomised cell line-combinations", x= "Pathway 1", y="Pathway 2") +
  scale_size_area( max_size = 4)


ggplot(full_Er_Es, aes(lib1_pathway, lib2_pathway)) + geom_point(aes(size=Es/Er, colour = (Es/Er>1)))+
  scale_colour_manual(values =c( "black","blue")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(size = "Fraction of cell line-combination pairs \nenriched for synergy", x= "Pathway 1", y="Pathway 2") +
  scale_size_area( max_size = 4)

```








