---
title: "Supplementary Figure 1"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(reshape2)
library(scales)
library(dplyr)
library(ggplot2)
library(superheat)
library(ggplot2)
library(dplyr)
library(magrittr)
library(knitr)
library(lubridate)
library(tidyr)
library(RColorBrewer)
library(reshape2)
library(knitr)
library(kableExtra)
library(scales)
library(ggpubr)
library(stats)
library(ggridges)
library(hexbin)
set.seed(13)
```


## Data import
Data files can be downloaded from Figshare (link tbc).
```{r}
sand1<-read.csv("Input/sandpiper-01_matrix_results.csv")
sand1$screen<-"Sandpiper_1"
g7<-read.csv("Input/gdsc-007_matrix_results.csv")
g7$screen<-"GDSC007"
g8<-read.csv("Input/gdsc-008_matrix_results.csv")
g8$screen<-"GDSC008"
g9<-read.csv("Input/gdsc-009_matrix_results.csv")
g9$screen<-"GDSC009"
g10<-read.csv("Input/gdsc-010_matrix_results.csv")
g10$screen<-"GDSC010"

data<-rbind(sand1, g7, g8, g9, g10)
```

## Figure S1A
Comparison with GDSC 2 monotherapies - identical protocol. GDSC2 data available for download from the [Genomics of Drug Sensitivity in Cancer website](https://www.cancerrxgene.org/downloads/drug_data?screening_set=GDSC2). 
```{r}
data_S1A_1<-data
data_S1A_2<-data
#Rename lib1 and lib2 ln IC50s as single agent ln IC50s
names(data_S1A_1)[17]<-"Sandpiper_ln_IC50"
names(data_S1A_2)[28]<-"Sandpiper_ln_IC50"
#Create compound variable of cell line ID and drug ID for matching to monotherapy data
data_S1A_1$test<-paste(data_S1A_1$SIDM, data_S1A_1$lib1_ID, sep="_")
data_S1A_2$test<-paste(data_S1A_2$SIDM, data_S1A_2$lib2_ID, sep="_")
data_S1A_1<-select(data_S1A_1, test, Sandpiper_ln_IC50)
data_S1A_2<-select(data_S1A_2, test, Sandpiper_ln_IC50)
data_S1A<-rbind(data_S1A_1, data_S1A_2)
#Aggregate replicates by mean
data_S1A<-aggregate(x = data_S1A$Sandpiper_ln_IC50,
                by = list(data_S1A$test),              
                FUN = mean)   
names(data_S1A)[2]<-"Sandpiper_ln_IC50"
names(data_S1A)[1]<-"test"

#Read in previously published single agent ln_IC50s
gdsc2<-read.csv("Input/GDSC2_fitted_dose_response_25Feb20.csv")
gdsc2<-select(gdsc2, SANGER_MODEL_ID, DRUG_ID, LN_IC50)
names(gdsc2)[3]<-"GDSC2_ln_IC50"
#Create compound variable of cell line ID and drug ID for matching to combination single agent ln IC50s
gdsc2$test<-paste(gdsc2$SANGER_MODEL_ID, gdsc2$DRUG_ID, sep="_")
gdsc2<-gdsc2[,-1:-2]
#Aggregate replicates by mean
gdsc2<-aggregate(x = gdsc2$GDSC2_ln_IC50,
                 by = list(gdsc2$test),              
                 FUN = mean)   
names(gdsc2)[2]<-"GDSC2_ln_IC50"
names(gdsc2)[1]<-"test"

all_S1A<-merge(data_S1A, gdsc2, by="test")

#correlation between combination study ln IC50s and previously-published ln IC50s
cor(all_S1A$Sandpiper_ln_IC50, all_S1A$GDSC2_ln_IC50)

plot(all_S1A$GDSC2_ln_IC50, all_S1A$Sandpiper_ln_IC50, xlab="ln(IC50) in GDSC2", ylab="ln(IC50) for single agents in this study", pch=20)
```


## Figure S1B
Comparison with GDSC 1 monotherapies - similar protocol, correlation of 74.30%. GDSC1 data available for download from the [Genomics of Drug Sensitivity in Cancer website](https://www.cancerrxgene.org/downloads/drug_data?screening_set=GDSC1). 
```{r}
#Read in previously published single agent ln_IC50s
gdsc1<-read.csv("Input/GDSC1_fitted_dose_response_24Jul22.csv")
gdsc1<-select(gdsc1, SANGER_MODEL_ID, DRUG_ID, LN_IC50)
names(gdsc1)[3]<-"GDSC1_ln_IC50"
#Create compound variable of cell line ID and drug ID for matching to combination single agent ln IC50s
gdsc1$test<-paste(gdsc1$SANGER_MODEL_ID, gdsc1$DRUG_ID, sep="_")
gdsc1<-gdsc1[,-1:-2]
#Aggregate replicates by mean
gdsc1<-aggregate(x = gdsc1$GDSC1_ln_IC50,
                 by = list(gdsc1$test),              
                 FUN = mean)   
names(gdsc1)[2]<-"GDSC1_ln_IC50"
names(gdsc1)[1]<-"test"

all_S1B<-merge(data_S1A, gdsc1, by="test")

#correlation between combination study ln IC50s and previously-published ln IC50s
cor(all_S1B$Sandpiper_ln_IC50, all_S1B$GDSC1_ln_IC50)

plot(all_S1B$GDSC1_ln_IC50, all_S1B$Sandpiper_ln_IC50, xlab="ln(IC50) in GDSC1", ylab="ln(IC50) for single agents in this study", pch=20)

#699 cell lines in common, two compounds in common
```

## Figure S1C

There are 13 drugs and 346 cell lines in common between the two screens. There is a correlation of 51.3%, where data is available. PRISM data was downloaded from [DepMap Portal Custom Download page](https://depmap.org/portal/download/custom/), with the following options selected: use all cell lines; use all genes/compounds; Drug sensitivity (PRISM Repurposing Primary Screen). 
```{r}
prism<-read.csv("Input/PRISM/Drug_sensitivity_(PRISM_Repurposing_Primary_Screen).csv")
row.names(prism)<-prism[,1]
common_drugs<-read.csv("PRISM_Sanger_mapping.csv")

NameList <- unique(common_drugs$PRISM_name)
idx <- match(NameList, names(prism))
NewDF <- prism[,idx] 
NewDF$BROAD_ID<-row.names(NewDF)
test<-melt(NewDF)
names(test)[2]<-"PRISM_name"
names(test)[3]<-"PRISM_ln_IC50"

test<-merge(test, common_drugs, by= "PRISM_name")
test<-test[,-1]
#Model list from Cell Model Passports for cell line ID dictionary
cmp<-read.csv("Input/PRISM/model_list_20230608.csv")
cmp<-select(cmp, "model_id", "BROAD_ID")
test<-merge(test, cmp, by= "BROAD_ID")
test<-test[,-1]
test$test<-paste(test$model_id, test$Sanger.ID, sep="_")
test<-test[,-2:-4]

prism_results<-aggregate(x = test$PRISM_ln_IC50,
                 by = list(test$test),              
                 FUN = mean)   
names(prism_results)[2]<-"PRISM_ln_IC50"
names(prism_results)[1]<-"test"


all_prism<-merge(data_S1A, prism_results, by="test")
all_prism<-all_prism[complete.cases(all_prism), ]

#correlation between combination study ln IC50s and previously-published ln IC50s
cor(all_prism$Sandpiper_ln_IC50, all_prism$PRISM_ln_IC50)

plot(all_prism$PRISM_ln_IC50, all_prism$Sandpiper_ln_IC50, xlab="ln(IC50) in PRISM", ylab="ln(IC50) for single agents in this study", pch=20)

```



## Figure S1H
```{r}
data_1_scatter<-data
data_2_scatter<-data
#Renaming lib1 and lib2 Emax as single agent Emax
names(data_1_scatter)[16]<-"Single_agent_Emax"
names(data_2_scatter)[27]<-"Single_agent_Emax"
data_1_scatter<-select(data_1_scatter, Single_agent_Emax, combo_MaxE, Bliss_matrix, HSA_matrix)
data_2_scatter<-select(data_2_scatter, Single_agent_Emax, combo_MaxE, Bliss_matrix, HSA_matrix)
data_scatter<-rbind(data_1_scatter, data_2_scatter)

plot(data_scatter$Bliss_matrix, data_scatter$HSA_matrix, xlab="Bliss excess", ylab="HSA excess", pch=20)
abline(h = c(0, 0.1), col=c("gray", "red"), lty=c(1,2), lwd=c(1, 3))
abline(v = c(0, 0.1), col=c("gray", "red"), lty=c(1,2), lwd=c(1, 3))
cor(data_scatter$Bliss_matrix, data_scatter$HSA_matrix)
```

## Figure S1I
```{r}
plot(data_scatter$Bliss_matrix, data_scatter$combo_MaxE, xlab="Bliss excess", ylab="Combo Emax", pch=20)
abline(v = c(0, 0.1), col=c("gray", "red"), lty=c(1,2), lwd=c(1, 3))
```


## Figure S1J
```{r}
plot(data_scatter$HSA_matrix, data_scatter$combo_MaxE, xlab="HSA excess", ylab="Combo Emax", pch=20)
abline(v = c(0, 0.1), col=c("gray", "red"), lty=c(1,2), lwd=c(1, 3))

```

## Figure S1K
```{r}
plot(data_scatter$Single_agent_Emax, data_scatter$combo_MaxE, xlab="Single agent Emax", ylab="Combo Emax", pch=20)


```

## Figure S1L
```{r}
plot(data_scatter$Single_agent_Emax, data_scatter$Bliss_matrix, xlab="Single agent Emax", ylab="Bliss excess", pch=20)
abline(h = c(0, 0.1), col=c("gray", "red"), lty=c(1,2), lwd=c(1, 3))
```

## Figure S1M
```{r}
a<-read.csv("Input/plate_qc_Sandpiper-01_13May22.csv")
b<-read.csv("Input/plate_qc_Sandpiper-02_13May22.csv")
data_S1L<-rbind(a,b)
data_S1L<-subset(data_S1L, data_S1L$QC_FLAG=="P")
data_S1L<-subset(data_S1L, data_S1L$RATIO_NC1_TO_PC1>=4)
data_S1L<-subset(data_S1L, data_S1L$RATIO_NC1_TO_PC2>=4)
data_S1L$label<-"DMSO, NC1"

#CV plot -  dashed line is plotted as per threshold
p <- ggplot(data_S1L, aes(x=CV_NC1)) + geom_boxplot() + scale_color_grey() + theme_classic() + coord_flip() + labs(x="Coefficient of variation (CV)", ylab="DMSO\n NC1") + theme(axis.ticks = element_blank(), axis.text.x=element_blank()) + geom_vline(xintercept=0.18, linetype="dashed",  color = "gray", size=2) + xlim(0,0.2)
p

median(data_S1L$CV_NC1)
range(data_S1L$CV_NC1)
```

## Figure S1N
```{r}
data_S1L_1<-data_S1L$Z_NC1_PC1
data_S1L_2<-data_S1L$Z_NC1_PC2
data_S1L_B<-data_S1L$Z_NC1_B

data_S1L_1<-as.data.frame(data_S1L_1)
data_S1L_2<-as.data.frame(data_S1L_2)
data_S1L_B<-as.data.frame(data_S1L_B)
names(data_S1L_1)[1]<-"Value"
names(data_S1L_2)[1]<-"Value"
names(data_S1L_B)[1]<-"Value"
data_S1L_1$type<-"MG-132 (PC1)"
data_S1L_2$type<-"Staurosporin (PC2)"
data_S1L_B$type<-"Blank (B)"

new_S1M<-rbind(data_S1L_1, data_S1L_2, data_S1L_B)

#Z score - remember to change dashed line according to threshold
p <- ggplot(new_S1M, aes(x=Value, y=type)) + geom_boxplot() + scale_color_grey() + theme_classic() + coord_flip() + labs(x="Z-factor", y="Variable")  + geom_vline(xintercept=0.3, linetype="dashed", color = "gray", size=2) + xlim(0,0.9)
p


pos<-rbind(data_S1L_1, data_S1L_2)

median(pos$Value)
max(pos$Value) - min(pos$Value)

```


##Figure S1O
```{r}
library(ggplot2)
library(dplyr)
data_s<-read.csv("Input/sandpiper-01_matrix_results.csv")
data_s$Screen<-"Sandpiper1"

data_7<-read.csv("Input/gdsc-007_matrix_results.csv")
data_7$Screen<-"GDSC007"
data_8<-read.csv("Input/gdsc-008_matrix_results.csv")
data_8$Screen<-"GDSC008"
data_9<-read.csv("Input/gdsc-009_matrix_results.csv")
data_9$Screen<-"GDSC009"
data_10<-read.csv("Input/gdsc-010_matrix_results.csv")
data_10$Screen<-"GDSC010b"
data_7plus<-rbind(data_7, data_8, data_9, data_10)

All.data<-rbind(data_s, data_7plus)


#write function to shorten p-values
format_pval <- function(pval){
  pval <- scales::pvalue(pval, accuracy= 0.001, add_p = TRUE)
  gsub(pattern = "(=|<)", replacement = " \\1 ", x = pval)
}


Correlation.BioReps_Average.Responses <- All.data %>%
  filter(CELL_LINE_NAME %in% c("A375", "HT-29", "PC-14", "U-2-OS", "SW620", "C32", "MHH-ES-1")) %>%
  left_join(All.data%>%
              #mutate(Date = as.Date(DATE_CREATED, format = "%d-%b-%y")) %>%
              select(BARCODE,DRUGSET_ID) %>%
              distinct()) %>%
  group_by(DRUGSET_ID,CELL_LINE_NAME, lib1_ID, lib2_ID, Screen) %>%
  summarise(combo_MaxE = mean(combo_MaxE),
            Bliss_matrix = mean(Bliss_matrix),
            Bliss_window = mean(Bliss_window),
            HSA_matrix = mean(HSA_matrix),
            HSA_window = mean(HSA_window)) %>%
  ungroup %>%
  melt(measure.vars = c("combo_MaxE", "Bliss_matrix", "Bliss_window", "HSA_matrix", "HSA_window")) %>%
  full_join(All.data %>%
              filter(CELL_LINE_NAME %in% c("A375", "HT-29", "PC-14", "U-2-OS", "SW620", "C32", "MHH-ES-1")) %>%
              left_join(All.data %>%
                          #mutate(Date = as.Date(DATE_CREATED, format = "%d-%b-%y")) %>%
                          select(BARCODE,DRUGSET_ID)%>%
                          distinct())  %>%
              group_by(DRUGSET_ID,CELL_LINE_NAME, lib1_ID, Screen) %>%
              summarise(lib1_MaxE = mean(lib1_MaxE),
                        lib1_IC50_ln = mean(lib1_IC50_ln)) %>%
              ungroup %>%
              melt(measure.vars = c("lib1_MaxE", "lib1_IC50_ln"))) %>%
  full_join(All.data %>%
              #Filter to control cell lines
              filter(CELL_LINE_NAME %in% c("A375", "HT-29", "PC-14", "U-2-OS", "SW620", "C32", "MHH-ES-1")) %>%
              left_join(All.data%>%
                          #mutate(Date = as.Date(DATE_CREATED, format = "%d-%b-%y")) %>%
                          select(BARCODE,DRUGSET_ID) %>%
                          distinct())  %>%
              group_by(DRUGSET_ID,CELL_LINE_NAME, lib2_ID, Screen) %>%
              summarise(lib2_MaxE = mean(lib2_MaxE),
                        lib2_IC50_ln = mean(lib2_IC50_ln)) %>%
              ungroup %>%
              melt(measure.vars = c("lib2_MaxE", "lib2_IC50_ln"))) %>%
  rename(Metric = variable) %>%
  #arrange(Date) %>%
  group_by(CELL_LINE_NAME, lib1_ID,  lib2_ID, Screen, Metric) %>%
  mutate(Replicate = 1:n()) %>%
  ungroup() %>%
  mutate(Reorder_Metric = case_when(Metric == "lib1_MaxE" ~ 1,
                                    Metric == "lib1_IC50_ln" ~ 2,
                                    Metric == "lib2_MaxE" ~ 3,
                                    Metric == "lib2_IC50_ln" ~ 4,
                                    Metric == "combo_MaxE" ~ 5,
                                    Metric == "Bliss_matrix" ~ 6,
                                    Metric == "Bliss_window" ~ 7,
                                    Metric == "HSA_matrix" ~ 8,
                                    Metric == "HSA_window" ~ 9),
         Metric = case_when(Metric == "lib1_MaxE" ~ "lib1_MaxE",
                            Metric == "lib1_IC50_ln" ~ "lib1_IC50_ln",
                            Metric == "lib2_MaxE" ~ "lib2_MaxE",
                            Metric == "lib2_IC50_ln" ~ "lib2_IC50_ln",
                            Metric == "combo_MaxE" ~ "combo_MaxE",
                            Metric == "Bliss_matrix" ~ "Bliss_matrix",
                            Metric == "Bliss_window" ~ "Bliss_window",
                            Metric == "HSA_matrix" ~ "HSA_matrix",
                            Metric == "HSA_window" ~ "HSA_window")) %>%
  arrange(Reorder_Metric) %>%
  mutate(Metric = factor(Metric, unique(Metric)))


# Calculate Pearson correlation coefficients for each bio rep pair
# There might be easier ways, but...
# full join the above data frame with itself to get all bio rep pairs
# Remove self pairs (e.g. BR1 with BR1)
# Remove duplicated pairs (e.g. BR1/BR2 = BR2/BR1)
# group bio rep pair & metric and calculate pearson correlation coefficient

Correlation.BioReps_Coefficients <- Correlation.BioReps_Average.Responses %>%
  full_join(Correlation.BioReps_Average.Responses %>%
              mutate(Screen = "All")) %>%
  rename(Value_1 = value,
         Replicate_1 = Replicate) %>%
  full_join(Correlation.BioReps_Average.Responses %>%
              full_join(Correlation.BioReps_Average.Responses %>%
                          mutate(Screen = "All"))%>%
              select(-DRUGSET_ID) %>%
              rename(Value_2 = value,
                     Replicate_2 = Replicate)) %>%
  filter(!Replicate_1 == Replicate_2) %>%
  group_by(CELL_LINE_NAME, lib1_ID, lib2_ID, Screen, Metric,
           BioRep_Pair = paste(pmin(Replicate_1, Replicate_2), pmax(Replicate_1, Replicate_2), sep = "_")) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  group_by(BioRep_Pair, Metric, Screen) %>%
  mutate(Correlation = cor(Value_1, Value_2, method = "pearson", use = "pairwise.complete.obs"),
         p_value = cor.test(Value_1, Value_2, method = "pearson")$p.value) %>%
  ungroup()



knitr::kable(Correlation.BioReps_Coefficients %>%
               group_by(Screen, Metric) %>%
               summarise(Median_Pearson = round(median(Correlation), digits = 3)) %>%
               ungroup() %>%
               spread(Screen, Median_Pearson), 
             align=c('l', 'l', 'l', 'c', 'c', 'c', 'c', 'c', 'c', 'c'), format="html")%>%
  kable_styling(bootstrap_options=c("stripped","hover","condensed","bordered"), full_width=F, position="left", font_size=12)



# Plot all pearson correlation coefficients by input metric, display as boxplot

P_QC_Correlations.key.metrics <- ggplot(Correlation.BioReps_Coefficients %>%
                                          filter(Screen == "All"),
                                        aes(x = Metric, y = Correlation)) +
  geom_boxplot() +
  theme_bw() +
  ylab("Pearson correlation coefficient") +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  theme(panel.grid = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 25, hjust = 1),
        axis.title.y = element_text(size = 10)) 

P_QC_Correlations.key.metrics



#pdf("S1O_BioReps_Corr of metrics_All data.pdf", width = 4, height = 3)
#print(P_QC_Correlations.key.metrics)
#dev.off()

```